---
title: "P8105 Homework 1"
author: "Ziyang Tian"
date: "2026-09-25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
```

# Problem 1

```{r load-penguins}
data("penguins", package = "palmerpenguins")
```

The penguins dataset has r nrow(penguins) rows and r ncol(penguins) columns. It includes three penguin species and measurements such as bill length, flipper length, and body mass. The mean flipper length is r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2) mm.

```{r penguin-scatterplot, fig.width=7, fig.height=5}
penguin_scatterplot <-
  ggplot(
    data = penguins,
    mapping = aes(
      x = bill_length_mm,
      y = flipper_length_mm,
      color = species
    )
  ) +
  geom_point(na.rm = TRUE) +
  labs(
    title = "Flipper Length vs. Bill Length",
    x = "Bill length (mm)",
    y = "Flipper length (mm)",
    color = "Species"
  ) +
  theme_minimal()

penguin_scatterplot
```


```{r save-scatterplot}
ggsave(
  filename = "penguins_scatterplot.png",
  plot = penguin_scatterplot,
  width = 7,
  height = 5,
  units = "in"
)
```


# Problem 2

```{r create-dataframe}
set.seed(8105)

homework_df <-
  tibble(
    normal_sample = rnorm(n = 10),
    above_zero = normal_sample > 0,
    character_values = letters[1:10],
    factor_values = factor(
      c("low", "medium", "high", "low", "medium",
        "high", "low", "medium", "high", "low"),
      levels = c("low", "medium", "high")
    )
  )

homework_df
```

The four variables have different types: `normal_sample` is numeric, `above_zero` is logical, `character_values` is character, and `factor_values` is a factor with the three levels low, medium, and high.

```{r calculate-means}
mean(pull(homework_df, normal_sample))
mean(pull(homework_df, above_zero))
mean(pull(homework_df, character_values))
mean(pull(homework_df, factor_values))
```

mean() works for the numeric and logical columns. For the logical column, R counts TRUE as 1 and FALSE as 0, so its mean is the proportion of values above zero. The character and factor columns return NA because they are categories rather than numbers.mean() works for the numeric and logical columns. For the logical column, R counts TRUE as 1 and FALSE as 0, so its mean is the proportion of values above zero. The character and factor columns return NA because they are categories rather than numbers.

```{r coercion, results="hide"}
as.numeric(pull(homework_df, above_zero))
as.numeric(pull(homework_df, character_values))
as.numeric(pull(homework_df, factor_values))
```

With as.numeric(), logical values become 1 and 0. The letters become NA because they cannot be read as numbers. The factor becomes 1, 2, or 3 according to its level order. These are category codes, so taking their mean would not have a useful interpretation.
